\vspace{-1cm} \bfseries\large Phosphatidylinositol 4,5-bisphosphate regulates cilium transition zone maturation in \textit{Drosophila melanogaster}

### 2018-03-13 latex/paper.tex
Edit to address critique from B. Trimble

    %
    %   Results - 3
    %
    \subsection{Hyperelongated transition zones exhibit signs of functional defects}
    We sought to examine whether TZ hyperelongation due to SigD expression
    affected TZ function.
    During cell elongation following meiosis, the TZ detaches from
    the basal body and migrates along the growing axoneme, maintaining a ciliary compartment
    at the distal-most {\raise.17ex\hbox{$\scriptstyle\sim$}}5$\mu$m where tubulin
    is presumably incorporated into the axoneme
    \citep{basiri2014migrating, fabian2012drosophila}.
    We found that
    TZs that assembled in \sigd{} cells are non-migratory despite the growth of the
    axoneme and the cell membrane, as shown by Unc and Cep290 localization
    (Figures 3A and 3B).
    We previously described ``comet-shaped'' localization of Unc-GFP in these cells,
    and we find that these hyperelongated Unc signals persist into cellular elongation
    despite elongation of the axoneme (Figure 3A, bottommost panel).

    In Drosophila and humans, basal bodies consist of microtubule triplets
    \citep{jana2016drosophila, lattao2017centrioles}
    while the axoneme contains microtubule doublets due to the obstruction of
    the C-tubule at the TZ \citep{gottardo2013cilium}.
    Consistent with a defect in the TZ barrier and the presence of
    microtubule triplets in axonemes in \sigd{} \citep{wei2008depletion},
    we observed that a subset of cilia (<5\%) in SigD-expressing cells
    contained puncta of Ana1, a core centriolar protein, at the tips
    of TZs marked with Cep290 (Figure 3C).
    We found that treatment of germ cells with Taxol as in \citep{riparbelli2013unique}
    could increase the penetrance of this phenotype (Figure 3D).
    Taxol treatment stabilizes microtubules and is thought to disrupt TZ maturation
    by inhibiting microtubule rearrangements \citep{riparbelli2012assembly}.
    Taxol treatment alone did not induce this phenotype in wild-type controls in our hands
    at varying concentrations ($p$ <0.01 at 5\% penetrance).
    Triplet microtubules within the axoneme are also observed in Taxol-treated cells
    \citep{riparbelli2013unique}, further supporting a relationship between
    \PIP{} and microtubule reorganization in TZ maturation.
    Asterless (homolog of CEP152), another centriolar protein,
    did not localize to TZ-distal puncta in \sigd{}
    \citep{dzhindzhev2010asterless, blachon2008drosophila} ($p$ <0.01)
    suggesting that these TZ-distal
    sites are not fully centriolar in protein composition.

@@

    Due to the involvement of exocyst in membrane trafficking at the plasma membrane,
    we examined whether cilium-associated membrane dynamics were affected in \sigd{}
    or \textit{onr} mutants similar to \textit{dila; cby} mutants
    \citep{vieillard2016transition}.
    We found that while TZs in \sigd{} and \textit{onr} cells anchored properly
    to the plasma membrane, they were unable to either establish or maintain
    membrane connections, and were rendered cytoplasmic (Figure 4A and B) similar to
    \textit{dila; cby} mutants.
    Although we were not able to determine the localization of Exo84 due to the
    lack of available reagents,
    we found that GFP-tagged Exo70 localized to centrioles (Figure 4D).
    Exo70 can directly bind to \PIP{} and mediates the recruitment of
    the exocyst complex to the plasma membrane \citep{he2007exo70}.
    Our results suggest that the exocyst may act downstream of \PIP{} to
    enhance cilium-membrane associations, and that TZ hyperelongation and loss of
    membrane-cilium association are genetically separable phenotypes.

@@

    The majority of \PIP{} at the plasma membrane is produced by PIPKIs
    \citep{balla2013phosphoinositides, hammond2012pi4p}
    and these seem to have a conserved role at cilia.
    In this study, we showed that mutation of the PIPKI Skittles (Sktl)
    induced hyperelongated TZs, similar to SigD expression.
    We were able rescue TZ hyperelongation by expression of full-length Sktl to
    a large degree.
    Interestingly, some cells showed cilium-autonomous suppression of lengths.
    The development of the Drosophila male germline occurs in syncytial ``cysts''
    of cells which share the same cytoplasm, and
    our results suggest that Skittles might function \textit{in situ}
    to regulate TZ maturation in these cells.
    In humans, \textit{PIPKI$\gamma$} is linked to lethal congenital contractural
    syndrome type 3 (LCCS3), which has been suggested to represent a ciliopathy.
    The recent discovery of the role of LCCS1-associated GLE1 protein in cilium
    function \citep{jao2017role} further supports this hypothesis.
    Thus, PIPKIs might represent novel ciliopathy-associated genes
    or genetic modifiers of disease.

    The gene \textit{EXOC8}, encoding the human Exo84, has also
    been linked to Joubert syndrome, a ciliopathy
    \citep{dixon2012exome}.
    Members of the exocyst complex, such as Sec10 and Sec6, have been shown
    to be important for cilium formation in cultured cell lines and zebrafish
    \citep{zuo2009exocyst, lobo2017exocyst, seixas2016arl13b}.
    The exocyst is known to bind to \PIP{}
    \citep{he2007exo70, zhang2008membrane}, and we previously showed that
    the \textit{onr} allele of Drosophila \textit{Exo84} phenocopied
    male germline defects observed in \sigd{} \citep{fabian2010phosphatidylinositol}.
    In this paper, we show that the \textit{onr} allele of the Drosophila Exo84
    phenocopies the loss of cilium-membrane contacts in \sigd{},
    but not TZ hyperelongation.
    Proper TZ function is required for membrane tethering, as evidenced by
    \textit{dila; cby} mutants, which contain abnormal TZs
    and show similar cytoplasmic cilia similar as \sigd{} \citep{vieillard2016transition}.
    Our results suggest that the exocyst is required for this process,
    possibly downstream of \PIP{}.

@@

    \textbf{Figure 3. Hyperelongated transition zones display functional defects.}
    \begin{enumerate}[label={(\Alph*)}, nolistsep]
    \item GFP-tagged Unc is unable to detach from the basal body and migrate in spermatids expressing \sigd{}.
    \item Cep290 is unable to detach and migrate from the basal body at the onset of axoneme assembly in \sigd{} spermatids.
    \item Structured illumination micrographs of \sigd{} cells show TZ-distal puncta of the centriolar protein Ana1.
    \item Treatment of \sigd{} cells with the microtubule stabilizing drug Taxol increases the penetrance of TZ-distal Ana1.
    \item Quantification of Cep290 lengths in Taxol-treated control and \sigd{} cells.
      % Inset shows distribution of regression coefficients (i.e. effects of Taxol and SigD on Cep290 length) from a fully Bayesian analysis.
    \end{enumerate}

>>>>

    %
    %   Results - 3
    %
    \subsection{Hyperelongated transition zones exhibit signs of functional defects}
    We sought to examine whether TZ hyperelongation due to SigD expression
    affected TZ function.
    During cell elongation following meiosis, the TZ detaches from
    the basal body and migrates along the growing axoneme, maintaining a ciliary compartment
    at the distal-most {\raise.17ex\hbox{$\scriptstyle\sim$}}5$\mu$m where tubulin
    is presumably incorporated into the axoneme
    \citep{basiri2014migrating, fabian2012drosophila}.
    We found that
    TZs that assembled in \sigd{} cells are non-migratory despite the growth of the
    axoneme (see Figure 1E) and the cell membrane, as shown by Unc and Cep290 localization
    (Figures 3A and 3B).
    We previously described ``comet-shaped'' localization of Unc-GFP in these cells,
    and we find that these hyperelongated Unc signals persist into cellular elongation
    despite elongation of the axoneme (Figure 3A, bottommost panel).

    In Drosophila and humans, basal bodies consist of microtubule triplets
    \citep{jana2016drosophila, lattao2017centrioles}
    while the axoneme contains microtubule doublets due to the obstruction of
    the C-tubule at the TZ \citep{gottardo2013cilium}.
    Consistent with a defect in the TZ barrier and the presence of
    microtubule triplets in axonemes in \sigd{} \citep{wei2008depletion},
    we observed that a subset of cilia (<5\%) in SigD-expressing cells
    contained puncta of Ana1, a core centriolar protein, at the tips
    of TZs marked with Cep290 (Figure 3C).
    We found that treatment of germ cells with Taxol as in \citep{riparbelli2013unique}
    could increase the penetrance of this phenotype from <5\% in untreated cells
    to \textgreater25\% in cells treated with 4 $\mu$M Taxol (Figure 3D)
    without affecting Cep290 length significantly (Figure 3E).
    Taxol treatment stabilizes microtubules and is thought to disrupt TZ maturation
    by inhibiting microtubule rearrangements \citep{riparbelli2012assembly}.
    Taxol treatment alone did not induce this phenotype in wild-type controls in our hands
    at varying concentrations ($p$ <0.01 at 5\% penetrance).
    Triplet microtubules within the axoneme are also observed in Taxol-treated cells
    \citep{riparbelli2013unique}, further supporting a relationship between
    \PIP{} and microtubule reorganization in TZ maturation.
    Asterless (homolog of CEP152), another centriolar protein,
    did not localize to TZ-distal puncta in \sigd{}
    \citep{dzhindzhev2010asterless, blachon2008drosophila} ($p$ <0.01)
    suggesting that these TZ-distal
    sites are not fully centriolar in protein composition.

@@

    Due to the involvement of exocyst in membrane trafficking at the plasma membrane,
    we examined whether cilium-associated membrane dynamics were affected in \sigd{}
    or \textit{onr} mutants similar to \textit{dila; cby} mutants
    \citep{vieillard2016transition}.
    We found that while TZs in \sigd{} and \textit{onr} cells anchored properly
    to the plasma membrane, they were unable to either establish or maintain
    membrane connections, and were rendered cytoplasmic (Figure 4A and B) similar to
    \textit{dila; cby} mutants.
    Although we were not able to determine the localization of Exo84 due to the
    lack of available reagents,
    we found that GFP-tagged Exo70 localized to centrioles (Figure 4D).
    Exo70 can directly bind to \PIP{} and mediates the recruitment of
    the exocyst complex to the plasma membrane \citep{he2007exo70}.
    Our results suggest that exocyst regulates
    cilium-membrane associations similar to \PIP{}, and that TZ hyperelongation and loss of
    membrane-cilium association are genetically separable phenotypes.

@@

    The majority of \PIP{} at the plasma membrane is produced by PIPKIs
    \citep{balla2013phosphoinositides, hammond2012pi4p}
    and these seem to have a conserved role at cilia.
    In this study, we showed that mutation of the PIPKI Skittles (Sktl)
    induced hyperelongated TZs, similar to SigD expression.
    We were able rescue TZ hyperelongation by expression of full-length Sktl to
    a large degree.
    Interestingly, some cells showed cilium-autonomous suppression of lengths.
    Since the development of the Drosophila male germline occurs in syncytial ``cysts''
    of cells which share the same cytoplasm,
    cilium-autonomous rescue suggests that Skittles might function \textit{in situ}
    to regulate TZ maturation in these cells.
    In humans, \textit{PIPKI$\gamma$} is linked to lethal congenital contractural
    syndrome type 3 (LCCS3), which has been suggested to represent a ciliopathy.
    The recent discovery of the role of LCCS1-associated GLE1 protein in cilium
    function \citep{jao2017role} further supports this hypothesis.
    This is further evidence that PIPKIs might represent ciliopathy-associated genes
    or genetic modifiers of disease, as previously hypothesized
    \citep{xu2016phosphatidylinositol}.

    In this study, we report two phenotypes associated with
    hyperelongated TZs which we postulate to be indicators of
    TZ dysfunction.
    Firstly, we show that hyperelongated TZs are unable to
    detach from the basal bodies at the onset of axoneme elongation
    in spermatids, unlike normal cells \citep{basiri2014migrating, fabian2012drosophila}.
    However, axonemes can assemble despite the lack of a functional
    TZ or membrane association in \sigd{} cells.
    As we previously reported, these axonemes have frequently aberrant
    ultrastructure \citep{wei2008depletion}.
    This is similar to \textit{cep290} \citep{basiri2014migrating}
    and \textit{dila}; \textit{cby} \citep{vieillard2016transition}
    mutants,
    both of which can assemble spermatid axonemes despite lacking
    functional TZs.
    We also report a novel phenotype of TZ-distal Ana1 localization
    in \sigd{} cells.
    In flies, Ana1 is important for centriole elongation
    \citep{saurya2016drosophila}.
    The human ortholog of Ana1 called CEP295 regulates
    centriole elongation by building the distal centriole and
    enhancing post-translational modification of the centriolar
    microtubules \citep{chang2016cep295}.
    The observation that Drosophila \textit{cep290} mutants,
    which lack a functional TZ diffusion barrier
    \citep{basiri2014migrating},
    contain hyperelongated centrioles suggests that the
    TZ normally suppresses centriole length.
    Our results of Ana1 being present at the tips of hyperelongated TZs
    without an increase in centriole length suggest that \sigd{} expression
    does not entirely disrupt the TZ diffusion barrier, but certain aspects of it,
    such as the inability to restrict C-tubules and Ana1.
    We found that
    Asterless (Asl), a peri-centriolar protein
    \citep{blachon2008drosophila}
    important for centrosome formation and centriole duplication, did not
    exhibit this TZ-distal localization.
    This may be due to many reasons, such as the different dynamics of Ana1 and Asl
    loading onto centrioles \citep{fu2016conserved, saurya2016drosophila} and the
    more peripheral nature of Asl within the centriole \citep{blachon2008drosophila}

    Members of the exocyst complex, such as Sec10 and Sec6, have been shown
    to be important for cilium formation in cultured cell lines and zebrafish
    \citep{zuo2009exocyst, lobo2017exocyst, seixas2016arl13b},
    but their precise roles in ciliogenesis are not well-understood.
    The subunits Sec3 and Exo70 regulate exocyst targeting to the plasma
    membrane through a direct interaction with \PIP{}
    \citep{he2007exo70, zhang2008membrane}.
    We previously showed that
    the \textit{onr} allele of Drosophila \textit{Exo84} phenocopied
    defects in cell polarity and elongation observed in \sigd{}
    \citep{fabian2010phosphatidylinositol}.
    In this paper, we show that the \textit{onr} allele of the Drosophila Exo84
    phenocopies the loss of cilium-membrane contacts in \sigd{} similar to
    \textit{dila; cby} mutants \citep{vieillard2016transition},
    but not TZ hyperelongation.
    Thus, TZ dysfunction is not a prerequisite for lack of
    cilium-plasma membrane association in male germ cells,
    and Exo84 uniquely regulates the latter process.
    This result is supported by the Drosophila \textit{cep290} mutant,
    which lacks a functional TZ but retains cilium-plasma membrane
    association \cite{basiri2014migrating}.
    Notably, \textit{EXOC8}, which encodes the human Exo84, has
    been linked to the ciliopathy Joubert syndrome
    \citep{dixon2012exome},
    and a similar process might underlie defects in humans with
    mutations in \textit{EXOC8}.

@@

    \textbf{Figure 3. Hyperelongated transition zones display functional defects.}
    \begin{enumerate}[label={(\Alph*)}, nolistsep]
    \item GFP-tagged Unc is unable to detach from the basal body and migrate in spermatids expressing \sigd{}.
      Cell elongation in spermatids is concomitant with elongation of the mitochondrial derivative (dark organelles in the phase contrast images).
    \item Cep290 is unable to detach and migrate from the basal body at the onset of axoneme assembly in \sigd{} spermatids.
    \item Structured illumination micrographs of \sigd{} cells show TZ-distal puncta of the centriolar protein Ana1.
    \item Treatment of \sigd{} cells with the microtubule stabilizing drug Taxol increases the penetrance of TZ-distal Ana1.
    \item Quantification of Cep290 lengths in Taxol-treated control and \sigd{} cells.
      % Inset shows distribution of regression coefficients (i.e. effects of Taxol and SigD on Cep290 length) from a fully Bayesian analysis.
    \end{enumerate}
